---
title: "Revenue Prediction Using Marketing Spend and Economic Indicators"
output: html_notebook
---

### **Project Objective**

Analyze and forecast the financial performance of a retail company by:

-   Forecasting future monthly revenue based on past trends and economic/marketing indicators.

-   Evaluating profitability, liquidity, and operational efficiency using financial ratios.

-   Simulating what-if scenarios (e.g., increased marketing spend).

-   Visualizing financial health and forecasts using professional dashboards.

**1. Monthly Forecast Dataset (2019–2023)**

-   **Columns**:

    -   `Month`: Monthly date

    -   `Revenue`: Total revenue with seasonality & trend

    -   `MarketingSpend`: Influenced by revenue + noise

    -   `UnemploymentRate`: Simulated macroeconomic factor

-   **Use cases**:

    -   Time series forecasting (ARIMA/ETS)

    -   Regression with lagged variables

    -   Scenario simulations (e.g., increase marketing)

### **Load necessary libraries and read dataset**

```{r}
install.packages("ggthemes")
# 
library(tidyverse)
library(lubridate)
library(forecast)
library(ggplot2)
library(tsibble)
library(scales)
library(ggthemes)
summary(forecast_data)
# Read the dataset
forecast_data <- read.csv("monthly_forecast_dataset.csv")



```

### Convert Month to Date

```{r}
forecast_data$Month <- as.Date(forecast_data$Month)



```

### Visualize Revenue Trend Over Time

```{r}

ggplot(forecast_data, aes(x = Month, y = Revenue)) +
  geom_line(color = "steelblue", size = 1) +
  labs(title = "Monthly Revenue Trend", x = "Month", y = "Revenue") +
  scale_y_continuous(labels = dollar_format()) +
  theme_minimal()



```

### Convert Revenue Column to a Time Series Object

```{r}
# Create a time series starting from Jan 2019 with monthly frequency
revenue_ts <- ts(forecast_data$Revenue, start = c(2019, 1), frequency = 12)

# Plot the time series to view pattern
autoplot(revenue_ts) +
  labs(title = "Revenue Time Series",
       x = "Year", y = "Revenue")
```

### Fit an ARIMA Forecasting Model

```{r}
# Automatically choose the best ARIMA model
model_arima <- auto.arima(revenue_ts)

# Show the model's details
summary(model_arima)

# Forecast for the next 6 months
forecast_arima <- forecast(model_arima, h = 6)

autoplot(forecast_arima) +
  labs(title = "Revenue Forecast for Next 6 Months",
       x = "Year", y = "Revenue") +
  scale_y_continuous(labels = dollar_format())
```

### Regression Model to Predict Revenue

```{r}
# Fit a linear regression model
model_reg <- lm(Revenue ~ MarketingSpend + UnemploymentRate, data = forecast_data)

# Show the results
summary(model_reg)

```

### Visualize Actual vs Predicted Revenue

```{r}
# Add predicted values to the original data
forecast_data$PredictedRevenue <- predict(model_reg, newdata = forecast_data)

# Plot actual vs predicted
ggplot(forecast_data, aes(x = Month)) +
  geom_line(aes(y = Revenue, color = "Actual Revenue"), size = 1) +
  geom_line(aes(y = PredictedRevenue, color = "Predicted Revenue"), size = 1, linetype = "dashed") +
  labs(title = "Actual vs Predicted Revenue",
       x = "Month", y = "Revenue") +
  scale_y_continuous(labels = dollar_format()) +
  scale_color_manual(values = c("Actual Revenue" = "steelblue", "Predicted Revenue" = "darkred")) +
  theme_minimal()
```
